### 2014-08-08 writeups/context-dependence.tex
More comments!
@petrelharp could you please look at the YYY comments, and decide
whether you like them or not? I think that heading in the direction
of more consistency will keep me sane.

For inference, we need to compute the $|\calS|^{m+2\ell} \times |\calS|^m$ matrix whose $(x,y)^\text{th}$ entry is $p_{m,\ell}(t,x,y)$,
for various $t$.
This matrix is a projection of the $|\calS|^{m+2\ell} \times |\calS|^{m+2\ell}$ matrix whose $(x,z)^\text{th}$ entry is
\begin{align}
    p_{m+2\ell}(x,z) = \left( e^{t G(m+2\ell)} \right)_{x,z} ,
\end{align}
where $G(m+2\ell)$ is the sparse $|\calS|^{m+2\ell} \times |\calS|^{m+2\ell}$ matrix defined in \eqref{eqn:G_defn}.
The ``projection'' we need just marginalizes over long patterns $z$ that match the shorter pattern $y$:
if we define the matrix $U$ in this way, so that $U_{zy}=1$ if $z_\ell^m=y$ and $U_{zy}=0$ otherwise,
then
\[
    p_{m,\ell}(t,x,y) = \sum_{z \in \calS^{m+2\ell}} \left( e^{t G(m+2\ell)} \right)_{xz} U_{zy} ,
\]
so in fact we don't need the entire matrix $e^{t G(m+2\ell)}$,
just the product of this matrix with each of the columns of $U$.

>>>>

For inference, we need to compute the $|\calS|^{m+2\ell} \times |\calS|^m$ matrix whose $(x,y)^\text{th}$ entry is $p_{m,\ell}(t,x,y)$,
for various $t$.
This matrix is a projection of the $|\calS|^{m+2\ell} \times |\calS|^{m+2\ell}$ matrix whose $(x,z)^\text{th}$ entry is
\begin{align}
    p_{m+2\ell}(x,z) = \left( e^{t G(m+2\ell)} \right)_{x,z} ,
\end{align}
where $G(m+2\ell)$ is the sparse $|\calS|^{m+2\ell} \times |\calS|^{m+2\ell}$ matrix defined in \eqref{eqn:G_defn}.
The ``projection'' we need just marginalizes over long patterns $z$ that match the shorter pattern $y$:
if we define the matrix $U$ in this way, so that $U_{zy}=1$ if $z_\ell^m=y$ and $U_{zy}=0$ otherwise,
then
\begin{align} \label{eqn:Tmer_trans}
    p_{m,\ell}(t,x,y) = \sum_{z \in \calS^{m+2\ell}} \left( e^{t G(m+2\ell)} \right)_{xz} U_{zy} ,
\end{align}
so in fact we don't need the entire matrix $e^{t G(m+2\ell)}$,
just the product of this matrix with each of the columns of $U$.
